---
title: "Clustering project final"
output: html_document
date: "2023-12-18"
---

::: {style="text-align:justify"}
Clustering is a method of assigning unlabeled data into relatively similar class groups. The idea is to categorize the data in such a way as to have distinctive differences between the groups while at the same time minimizing the differences between the objects inside the groups. There are many algorithms that allow clustering of the data, each with unique properties and methodology to suit different data types.

In this project, I will attempt to cluster the wine data by its properties and analyze the outcome.
:::

## Libraries

```{r}
library(cluster)
library(ClusterR)
library(stats)
library(factoextra)
library(gridExtra)
library(dplyr)
library(huxtable)
library(dplyr)
library(NbClust)
library(ggplot2)
```

::: {style="text-align:justify"}
To start with, I will save my wine data into the wine variable.
:::

```{r}
wine <- read.csv("winequalityN.csv")
```

::: {style="text-align:justify"}
Inspecting the data
:::

```{r}
str(wine)
head(wine)
```

::: {style="text-align:justify"}
Any missing values?
:::

```{r}
which(colSums(is.na(wine))>0)

```

::: {style="text-align:justify"}
Few missing values, thus I will just simply get rid of the rows with missing values.
:::

```{r}
wine<-na.omit(wine)

which(colSums(is.na(wine))>0)
```

::: {style="text-align:justify"}
Summary stat
:::

```{r}
summary(wine)
```

::: {style="text-align:justify"}
Since I want to cluster only the properties, I will get rid of type and quality columns.
:::

```{r}
properties <- wine[c(2:12)]
```

::: {style="text-align:justify"}
Also, because the variables are not of the same scale, scaling is necessary.
:::

```{r}
properties2 <- as.data.frame(lapply(properties, scale))

```

::: {style="text-align:justify"}
Is the dataset clusterable at all?
:::

```{r}
get_clust_tendency(properties2, 2, graph=FALSE, gradient=list(low="red", mid="white", high="blue"), seed = 123)
```

::: {style="text-align:justify"}
According to the documentation of get_clust_tendency, if the Hopkins stat is close to 1 (far above 0.5) then the dataset is clusterable. Since the Hopkins stat for my dataset is roughly 0.956 I can conclude that the dataset is highly clusterable.

The choice of the clustering method plays a crucial role in the quality of the analysis. Since my dataset is of moderate size, I believe the K-means method would be appropriate here. Furthermore, K-means is a computationally efficient algorithm with great interpretational powers for data analysis and gaining insights into the structure of the data.

What is the optimal number of clusters?
:::

```{r}
opt2<-NbClust(properties2, distance="euclidean", min.nc=2, max.nc=8, method="complete", index="ch") 

opt2$Best.nc
```

::: {style="text-align:justify"}
Based on the index given by NbClust, the optimal number of clusters would be 3.
:::

```{r}
opt1<-Optimal_Clusters_KMeans(properties2, max_clusters=10, plot_clusters=TRUE, criterion="silhouette")
```

::: {style="text-align:justify"}
The silhouette width confirms the results of NbClust, indicating 3 clusters as the most optimal.
:::

# K-means clustering

```{r}
cluster_kmeans <- eclust(properties2, k=3, FUNcluster="kmeans", hc_metric="euclidean", graph=FALSE) 
z <- fviz_silhouette(cluster_kmeans) 
x <- fviz_cluster(cluster_kmeans, data = White_wine, elipse.type = "convex") + theme_minimal() 
grid.arrange(z, x, ncol=2)
```

::: {style="text-align:justify"}
The results of clustering are not exceptionally high. The average silhouette value is 0.24. There are some outliers visible in particular in cluster 1. However, I still believe the result represents reasonable separation between clusters which will allow a meaningful interpretation.

Assigning the clusters to the original wine dataset:
:::

```{r}
wine$Cluster <- cluster_kmeans$cluster
```

# Analysis of the clusters

::: {style="text-align:justify"}
Clusters' centers
:::

```{r}
cc <- as_hux(data.frame(t(cluster_kmeans$centers)), add_rownames = TRUE)
colnames(cc) <- c("Feature", paste("Cluster", 1:3)) 
cc <- add_colnames(cc) 
cc
```

::: {style="text-align:justify"}
Size of clusters
:::

```{r}
cluster_assigement <- cluster_kmeans$cluster 
cluster_size <- table(cluster_assigement) 
cluster_size
```

::: {style="text-align:justify"}
Frequency
:::

```{r}
ggplot(wine, aes(x = factor(Cluster), fill = type)) + geom_bar(position = "stack") + labs(title = "Frequency Distribution of Categorical Variable by Cluster", x = "Cluster", y = "Count")
```

::: {style="text-align:justify"}
As seen on the graph, there is pretty clear division between red and white wines.
:::

::: {style="text-align:justify"}
Mean values
:::

```{r}
meanvalues1 <- wine %>% group_by(Cluster) %>% summarize( Mean_Quality = mean(quality), Mean_Fixed.acidity = mean(fixed.acidity), Mean_Volatile.acidity = mean(volatile.acidity), Mean_Citric.acid = mean(citric.acid), )

meanvalues2 <- wine %>% group_by(Cluster) %>% summarize( Mean_Residual.sugar = mean(residual.sugar), Mean_Chlorides = mean(chlorides), Mean_Free.sulfur.dioxide = mean(free.sulfur.dioxide), Mean_Total.sulfur.dioxide = mean(total.sulfur.dioxide), )

meanvalues3 <- wine %>% group_by(Cluster) %>% summarize( Mean_Density = mean(density), Mean_PH = mean(pH), Mean_Sulphates = mean(sulphates), Mean_Alcochol = mean(alcohol) ) 

meanvalues1 
meanvalues2 
meanvalues3
```

::: {style="text-align:justify"}
Analyzing the mean values of variables in each cluster uncovers a few interesting details. First of all, the lowest mean quality value appears in cluster 3, which consists mainly of red wines, while the highest appears in cluster 2, which is mainly made of white wines. However, it doesn't suggest that red wines are generally worse than white wines. If we compare cluster 1, which also consists mainly of white wines, with cluster 3, we can conclude that the mean quality values in these clusters are pretty similar. Second of all, I can say that generally, the acidity is higher in red wines, while the sulfur dioxide is higher in white wines. Interestingly enough, the value of the mean residual sugar skyrocketed in cluster 1, very likely due to some outliers.
:::

::: {style="text-align:justify"}
Boxplots of variables with significant differences in means between clusters.
:::

```{r}
data_with_clusters1 <- data.frame(volatile.acidity = wine$volatile.acidity, Cluster = factor(cluster_assigement))

a<-ggplot(data_with_clusters1, aes(x = Cluster, y = volatile.acidity, fill = Cluster)) + geom_boxplot() + labs(title = "Volatile acidity by Cluster", x = "Cluster", y = "Volatile acidity") + theme_minimal()

data_with_clusters2 <- data.frame(residual.sugar = wine$residual.sugar, Cluster = factor(cluster_assigement))

b<-ggplot(data_with_clusters2, aes(x = Cluster, y = residual.sugar, fill = Cluster)) + geom_boxplot() + labs(title = "Residual sugar by Cluster", x = "Cluster", y = "Residual sugar") + theme_minimal()

data_with_clusters3 <- data.frame(free.sulfur.dioxide = wine$free.sulfur.dioxide, Cluster = factor(cluster_assigement))

c<-ggplot(data_with_clusters3, aes(x = Cluster, y = free.sulfur.dioxide, fill = Cluster)) + geom_boxplot() + labs(title = "Free sulfur dioxide by Cluster", x = "Cluster", y = "Free sulfur dioxide") + theme_minimal()

data_with_clusters4 <- data.frame(total.sulfur.dioxide = wine$total.sulfur.dioxide, Cluster = factor(cluster_assigement))

d<-ggplot(data_with_clusters4, aes(x = Cluster, y = total.sulfur.dioxide, fill = Cluster)) + geom_boxplot() + labs(title = "Total sulfur dioxide by Cluster", x = "Cluster", y = "Total sulfur dioxide") + theme_minimal()

grid.arrange(a, b, c, d, ncol=2)
```

::: {style="text-align:justify"}
As expected, there are some outliers in cluster 1 for residual sugar, however, the interquartile range in that cluster is still higher in comparison to clusters 2 and 3. The free sulfur dioxide and total sulfur dioxide follow the same pattern of order, though it seems there is a bit bigger variation between the clusters for the total sulfur dioxide variable. Also some outliers mainly in cluster 1. The bar plot for volatile acidity shows little difference between clusters 1 and 2 while cluster 3 ranges much higher.
:::
